---
title: "Pre processed Thesis Code"
output:
  pdf_document: default
  html_document: default
date: "2023-10-24"
---

```{r setup, include=FALSE,cache=TRUE}
knitr::opts_chunk$set(echo = FALSE, tidy.opts = list(width.cutoff = 40),
tidy = TRUE)
pacman::p_load(tidyverse, plyr, dplyr, pacman,
tableone, ggcorrplot, GGally, knitr,
ggpubr, magrittr, broom, formatR, readxl,
tinytex, Metrics, lmtest, caret, glmtoolbox,
predtools, pROC, ggsurvfit, survival,tableone,
lubridate, gtsummary, tidycmprsk, survminer,gtools,
condSURV,naniar,kableExtra,tinytex)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


```{r,cache=TRUE}
library(nhanesA)
#nhanesCodebook(nh, colname, dxa = FALSE)
# first, load in all the lead tables
# then merge all into one dataset
# repeat for all the other vars
lead_1999 <- nhanes("LAB06")
lead_2001 <- nhanes("L06_2_B")
lead_2003 <- nhanes("L06BMT_C")
lead_2005 <- nhanes("PbCd_D")
lead_2007 <- nhanes("PbCd_E")
lead_2009 <- nhanes("PbCd_F")
lead_2011 <- nhanes("PbCd_G")
lead_2013 <- nhanes("PbCd_H")
lead_2015 <- nhanes("PbCd_I")
lead_2017 <- nhanes("PbCd_J")
lead_1999$YEAR <- 1999
lead_2001$YEAR <- 2001
lead_2003$YEAR <- 2003
lead_2005$YEAR <- 2005
lead_2007$YEAR <- 2007
lead_2009$YEAR <- 2009
lead_2011$YEAR <- 2011
lead_2013$YEAR <- 2013
lead_2015$YEAR <- 2015
lead_2017$YEAR <- 2017
lead_2001$LBXBPB <- lead_2001$LB2BPB
lead_2001<-lead_2001 %>%
  select(-c(LB2BPB))
lead_df <- list(lead_1999, lead_2001, lead_2003,
lead_2005, lead_2007, lead_2009, lead_2011,
lead_2013, lead_2015, lead_2017)
lead_df_1 <- join_all(lead_df, type = "full")
lead_df_final<-lead_df_1 %>%
  select(SEQN, YEAR, LBXBPB)
#print(lead_df_final)
```
The data displays the Lead by blood level (ug/dL) from the years 1999 to 2017. The dataset has 3 columns which include the sequence number (PATIENT ID), Year (year of measurement), and LBXBPB the measurement of lead by blood level. There are 76,213 rows in the original combined set. There are missing values observed in the measurements of LBXBPB
```{r,cache=TRUE}
# k. so. they have hypertension if
# HTN=yes, if medication = yes, and if
# systolic BP >= 130 mmHg or diastolic
# BP >=80 mmHg going to need to make a
# new column of final_HTN at the end of
# the full dataframe after selecting
# htn, medication, and both blood
# pressures
htn_1999 <- nhanes("BPQ")
htn_2001 <- nhanes("BPQ_B")
htn_2003 <- nhanes("BPQ_C")
htn_2005 <- nhanes("BPQ_D")
htn_2007 <- nhanes("BPQ_E")
htn_2009 <- nhanes("BPQ_F")
htn_2011 <- nhanes("BPQ_G")
htn_2013 <- nhanes("BPQ_H")
htn_2015 <- nhanes("BPQ_I")
htn_2017 <- nhanes("BPQ_J")
htn_1999$YEAR <- 1999
htn_2001$YEAR <- 2001
htn_2003$YEAR <- 2003
htn_2005$YEAR <- 2005
htn_2007$YEAR <- 2007
htn_2009$YEAR <- 2009
htn_2011$YEAR <- 2011
htn_2013$YEAR <- 2013
htn_2015$YEAR <- 2015
htn_2017$YEAR <- 2017

htn_df <- list(htn_1999, htn_2001, htn_2003,
htn_2005, htn_2007, htn_2009, htn_2011,
htn_2013, htn_2015, htn_2017)


htn_df_1 <- join_all(htn_df, type = "full")
htn_df_final <- htn_df_1 %>%
select(SEQN, YEAR, BPQ020, BPQ040A)
#print(htn_df_final)

# repeat process: load in all the BP
# tables then merge all into one
# dataset
bp_1999 <- nhanes("BPX")
bp_2001 <- nhanes("BPX_B")
bp_2003 <- nhanes("BPX_C")
bp_2005 <- nhanes("BPX_D")
bp_2007 <- nhanes("BPX_E")
bp_2009 <- nhanes("BPX_F")
bp_2011 <- nhanes("BPX_G")
bp_2013 <- nhanes("BPX_H")
bp_2015 <- nhanes("BPX_I")
bp_2017 <- nhanes("BPX_J")

bp_1999$YEAR <- 1999
bp_2001$YEAR <- 2001
bp_2003$YEAR <- 2003
bp_2005$YEAR <- 2005
bp_2007$YEAR <- 2007
bp_2009$YEAR <- 2009
bp_2011$YEAR <- 2011
bp_2013$YEAR <- 2013
bp_2015$YEAR <- 2015
bp_2017$YEAR <- 2017

bp_df <- list(bp_1999, bp_2001, bp_2003,
bp_2005, bp_2007, bp_2009, bp_2011, bp_2013,
bp_2015, bp_2017)

bp_df_1<-join_all(bp_df, type="full")
str(bp_df_1)

bp_df_final <- bp_df_1 %>%
  select(SEQN, YEAR, BPXSY1, BPXSY2, BPXSY3,BPXSY4, BPXDI1, BPXDI2, BPXDI3, BPXDI4)
#print(bp_df_final)
```
The hypertension data displays two binary measurements  from the years 1999 to 2017. The dataset has 4 columns which include the sequence number (PATIENT ID), Year (year of measurement), BPQ020 a binary measurement answering if patient has "ever been told by a doctor or other health professional that {you/s/he} had hypertension, also called high blood pressure?", and BPQ040 another binary measurement answering if patient "Because of {your/SP's} (high blood pressure/hypertension), {have you/has s/he} ever been told to . . . take prescribed medicine?". The indicators for both binary variables is as follows (1=Yes, 2=No). There are 63,592 rows in the  combined set. There are multiple missing values observed in the measurements of BPQ040A we might want to  consider that they may have never been asked.


The blood pressure data displays 8 numerical measurements from the years 1999 to 2017 for systolic and diastolic blood pressure points at different time points within the year. The dataset has 10 columns which include the sequence number (PATIENT ID), Year (year of measurement), BPXSY1,BPXSY2,BPXSY3, BPXSI4 which are readings for systolic blood pressure. The numbers at the end of the variables indicate the reading number and usually the fourth is only necessary for extreme cases.It's measurement units are mm Hg. BPXDI1,BPXDIY2,BPXDI3, BPXDI4  which are readings for diastolic blood pressure. The numbers at the end of the variables indicate the reading number and usually the fourth is only necessary for extreme cases.It's measurement units are mm Hg. There are 96,766 rows in the  combined set. There are multiple missing values observed in the measurements in all measurements for  systolic and diastolic and bp. 

```{r,cache=TRUE}
# repeat process: load in all the demo
# tables then merge all into one
# dataset
demo_1999 <- nhanes("DEMO")
demo_2001 <- nhanes("DEMO_B")
demo_2003 <- nhanes("DEMO_C")
demo_2005 <- nhanes("DEMO_D")
demo_2007 <- nhanes("DEMO_E")
demo_2009 <- nhanes("DEMO_F")
demo_2011 <- nhanes("DEMO_G")
demo_2013 <- nhanes("DEMO_H")
demo_2015 <- nhanes("DEMO_I")
demo_2017 <- nhanes("DEMO_J")

demo_1999$YEAR <- 1999
demo_2001$YEAR <- 2001
demo_2003$YEAR <- 2003
demo_2005$YEAR <- 2005
demo_2007$YEAR <- 2007
demo_2009$YEAR <- 2009
demo_2011$YEAR <- 2011
demo_2013$YEAR <- 2013
demo_2015$YEAR <- 2015
demo_2017$YEAR <- 2017

demo_df<- list(demo_1999, demo_2001, demo_2003,
demo_2005, demo_2007, demo_2009, demo_2011,
demo_2013, demo_2015, demo_2017)
demo_df_1<-join_all(demo_df, type ="full")
demo_df_final<- demo_df_1 %>%
  select(SEQN, YEAR, DMDHRAGE, RIDRETH1,
INDFMPIR, DMDEDUC2, RIAGENDR)

#print(demo_df_final)

```
The blood pressure data displays 5 numerical measurements from the years 1999 to 2017 for patient demographics including the patient age, patient ethnicity/race, ratio of income for patient, education level and gender. Refer to the codebook in excel sheet to see the measurements of each variable. The demographics table has 101,316 rows and 7 columns. There seems to be missingness for measurements in income ratio and education level.  


The data displays two binary measurements  from the years 1999 to 2017. The datasethas 4 columns which include the sequence nuber (PATIENT ID), Year (year of measurement), BPQ020 a binary measurement answering if patient has "ever been told by a doctor or other health professional that {you/s/he} had hypertension, also called high blood pressure?", and BPQ040 another binary measurement answering if patient "Because of {your/SP's} (high blood pressure/hypertension), {have you/has s/he} ever been told to . . . take prescribed medicine?". The indicators for both binary variables is as follows (1=Yes, 2=No). There are 63,592 rows in the  combined set. There are multiple missing values observed in the measurements of BPQ040A we might want to  consider that they may have never been asked.

```{r,cache=TRUE}
# repeat process: load in all the BMI
# tables then merge all into one
# datasetv
bmi_1999 <- nhanes("BMX")
bmi_2001 <- nhanes("BMX_B")
bmi_2003 <- nhanes("BMX_C")
bmi_2005 <- nhanes("BMX_D")
bmi_2007 <- nhanes("BMX_E")
bmi_2009 <- nhanes("BMX_F")
bmi_2011 <- nhanes("BMX_G")
bmi_2013 <- nhanes("BMX_H")
bmi_2015 <- nhanes("BMX_I")
bmi_2017 <- nhanes("BMX_J")

bmi_1999$YEAR <- 1999
bmi_2001$YEAR <- 2001
bmi_2003$YEAR <- 2003
bmi_2005$YEAR <- 2005
bmi_2007$YEAR <- 2007
bmi_2009$YEAR <- 2009
bmi_2011$YEAR <- 2011
bmi_2013$YEAR <- 2013
bmi_2015$YEAR <- 2015
bmi_2017$YEAR <- 2017

bmi_df<-list(bmi_1999, bmi_2001, bmi_2003,
bmi_2005, bmi_2007, bmi_2009, bmi_2011, bmi_2013, bmi_2015, bmi_2017)
bmi_df_1<-join_all(bmi_df, type ="full")
bmi_df_final <- bmi_df_1 %>%
  select(SEQN, YEAR, BMXBMI)
#print(bmi_df_final)
```

```{r,cache=TRUE}
# repeat process: load in all the
# Alcohol tables then merge all into
# one dataset
alc_1999 <- nhanes("ALQ")
alc_2001 <- nhanes("ALQ_B")
alc_2003 <- nhanes("ALQ_C")
alc_2005 <- nhanes("ALQ_D")
alc_2007 <- nhanes("ALQ_E")
alc_2009 <- nhanes("ALQ_F")
alc_2011 <- nhanes("ALQ_G")
alc_2013 <- nhanes("ALQ_H")
alc_2015 <- nhanes("ALQ_I")
alc_2017 <- nhanes("ALQ_J")

alc_1999$YEAR <- 1999
alc_2001$YEAR <- 2001
alc_2003$YEAR <- 2003
alc_2005$YEAR <- 2005
alc_2007$YEAR <- 2007
alc_2009$YEAR <- 2009
alc_2011$YEAR <- 2011
alc_2013$YEAR <- 2013
alc_2015$YEAR <- 2015
alc_2017$YEAR <- 2017

alc_df<-list(alc_1999, alc_2001, alc_2003,
alc_2005, alc_2007, alc_2009, alc_2011,
alc_2013, alc_2015, alc_2017)
alc_df_1<-join_all(alc_df, type="full")
alc_df_final<- alc_df_1 %>%
  select(SEQN, YEAR, ALQ120Q)
#print(alc_df_final)

# repeat process: load in all the
# Smoking tables then merge all into
# one dataset
smok_1999 <- nhanes("SMQ")
smok_2001 <- nhanes("SMQ_B")
smok_2003 <- nhanes("SMQ_C")
smok_2005 <- nhanes("SMQ_D")
smok_2007 <- nhanes("SMQ_E")
smok_2009 <- nhanes("SMQ_F")
smok_2011 <- nhanes("SMQ_G")
smok_2013 <- nhanes("SMQ_H")
smok_2015 <- nhanes("SMQ_I")
smok_2017 <- nhanes("SMQ_J")

smok_1999$YEAR <- 1999
smok_2001$YEAR <- 2001
smok_2003$YEAR <- 2003
smok_2005$YEAR <- 2005
smok_2007$YEAR <- 2007
smok_2009$YEAR <- 2009
smok_2011$YEAR <- 2011
smok_2013$YEAR <- 2013
smok_2015$YEAR <- 2015
smok_2017$YEAR <- 2017

smok_df<-list(smok_1999, smok_2001, smok_2003,
smok_2005, smok_2007, smok_2009, smok_2011,
smok_2013, smok_2015, smok_2017)
smok_df_1 <- join_all(smok_df, type= "full")
smok_df_final<-smok_df_1 %>%
  select(SEQN, YEAR, SMQ040, SMQ020)
#print(smok_df_final)
```

```{r,cache=TRUE}
# repeat process: load in all the
# Insurance tables then merge all into
# one dataset
ins_1999 <- nhanes("HIQ")
ins_2001 <- nhanes("HIQ_B")
ins_2003 <- nhanes("HIQ_C")
ins_2005 <- nhanes("HIQ_D")
ins_2007 <- nhanes("HIQ_E")
ins_2009 <- nhanes("HIQ_F")
ins_2011 <- nhanes("HIQ_G")
ins_2013 <- nhanes("HIQ_H")
ins_2015 <- nhanes("HIQ_I")
ins_2017 <- nhanes("HIQ_J")

ins_1999$YEAR <- 1999
ins_2001$YEAR <- 2001
ins_2003$YEAR <- 2003
ins_2005$YEAR <- 2005
ins_2007$YEAR <- 2007
ins_2009$YEAR <- 2009
ins_2011$YEAR <- 2011
ins_2013$YEAR <- 2013
ins_2015$YEAR <- 2015
ins_2017$YEAR <- 2017

ins_1999$HIQ011 <- ins_1999$HID010
ins_1999 <- ins_1999 %>%
  select(-c(HID010))
ins_2001$HIQ011 <- ins_2001$HID010
ins_2001 <- ins_2001 %>%
  select(-c(HID010))
ins_2003$HIQ011 <- ins_2003$HID010
ins_2003 <- ins_2003 %>%
  select(-c(HID010))
ins_df <- list(ins_1999, ins_2001, ins_2003,
ins_2005, ins_2007, ins_2009, ins_2011,
ins_2013, ins_2015, ins_2017)
ins_df_1 <- join_all(ins_df, type = "full")
ins_df_final <- ins_df_1 %>%
select(SEQN, YEAR, HIQ011)
#print(ins_df_final)
```

```{r, cache=TRUE}
# join all the datasets togther
nhanes_list<- list(lead_df_final, htn_df_final,
bp_df_final, demo_df_final, bmi_df_final,
alc_df_final, smok_df_final, ins_df_final)

nhanes_df<- join_all(nhanes_list, by = "SEQN", type = "full")
#print(nhanes_df)

# obtain overages for Systolic and
# Diastolic BP
nhanes_df$SYSAVG = with(nhanes_df, ifelse(!is.na(nhanes_df$BPXSY2) |
!is.na(nhanes_df$BPXSY3) | !is.na(nhanes_df$BPXSY4),
rowMeans(nhanes_df[, c("BPXSY2", "BPXSY3",
"BPXSY4")], na.rm = TRUE), nhanes_df$BPXSY1))

nhanes_df$DIAAVG = with(nhanes_df,ifelse(!is.na(nhanes_df$BPXDI2) |
!is.na(nhanes_df$BPXDI3) | !is.na(nhanes_df$BPXDI4),
rowMeans(nhanes_df[, c("BPXDI2", "BPXDI3",
"BPXDI4")], na.rm = TRUE), nhanes_df$BPXDI1))
head(nhanes_df)

# if BPQ020= 1 OR BPXSAR >= 130 mmHg OR
# BPXDAR >=80 mmHg, then htn=yes
nhanes_df$HTN <- ifelse((nhanes_df$BPQ020 ==
1) | (nhanes_df$SYSAVG >= 130) | (nhanes_df$DIAAVG >=
80), 1, 0)
head(nhanes_df)
str(nhanes_df)
# clean and merge the smok020 and
# smok040 variables to mimic original
# paper
nhanes_df$smokingstatus = with(nhanes_df,
ifelse(is.na(nhanes_df$SMQ020), NA, ifelse(nhanes_df$SMQ020 ==
2, 1, ifelse(nhanes_df$SMQ040 ==
1 | nhanes_df$SMQ040 == 2, 2, ifelse(nhanes_df$SMQ040 ==
3 & nhanes_df$SMQ020 == 1, 3, NA)))))

head(nhanes_df)
str(nhanes_df)
# drop all rows with variables (see
# list) that are na
library(tidyr)
miss_var_summary_result<-miss_var_summary(nhanes_df)
kable(miss_var_summary_result, caption = "Missing Value Summary")


nhanes_df1 <- nhanes_df %>%
  drop_na(LBXBPB, SYSAVG, DIAAVG)
str(nhanes_df1)

miss_var_summary_result1<-miss_var_summary(nhanes_df1)
kable(miss_var_summary_result1, caption = "Missing Value Summary")

nhanes_df2 <- nhanes_df1 %>%
  drop_na(DMDEDUC2, INDFMPIR, BMXBMI, ALQ120Q,HIQ011, smokingstatus)
# SMOKING variable drops a lot of
# variables. Insurance variable only
# drops two people (include insurance)
nhanes_df_final<- subset(nhanes_df2, DMDHRAGE >20)
str(nhanes_df_final)
nhanes_df_final <- nhanes_df_final %>%
  mutate(ALCSTATUS = ifelse(ALQ120Q ==0, "0", "1"))
str(nhanes_df_final)
```

```{r, cache=TRUE}
# factors
nhanes_df_final$HTN<- factor(nhanes_df_final$HTN)
nhanes_df_final$DMDEDUC2 <- factor(nhanes_df_final$DMDEDUC2)
nhanes_df_final$RIAGENDR <- factor(nhanes_df_final$RIAGENDR)
nhanes_df_final$RIDRETH1 <- factor(nhanes_df_final$RIDRETH1)
nhanes_df_final$smokingstatus <- factor(nhanes_df_final$smokingstatus)


nhanes_df_final<- nhanes_df_final[!nhanes_df_final$HIQ011 =="7", ] #dropping dont know and refused values, due to small sample size of these values (23) 
nhanes_df_final<- nhanes_df_final[!nhanes_df_final$HIQ011 =="9", ]


# rename the variables/columns and
# factor health insurance coverage
colnames(nhanes_df_final)[3] = "Blood Lead Level"
colnames(nhanes_df_final)[4] = "Hypertension Diagnosis"
colnames(nhanes_df_final)[14] = "Age"
colnames(nhanes_df_final)[15] = "Race/Ethnicity"
colnames(nhanes_df_final)[16] = "Ratio of Family Income to Poverty Level"
colnames(nhanes_df_final)[17] = "Education level"
colnames(nhanes_df_final)[18] = "Sex"
colnames(nhanes_df_final)[19] = "BMI"
colnames(nhanes_df_final)[20] = "Number of Drinks/Year"
colnames(nhanes_df_final)[23] = "Health Insurance Coverage"
colnames(nhanes_df_final)[24] = "Average Systolic Blood Pressure"
colnames(nhanes_df_final)[25] = "Average Diastolic Blood Pressure"
colnames(nhanes_df_final)[26] = "Hypertension Status"
colnames(nhanes_df_final)[27] = "Smoking Status"
colnames(nhanes_df_final)[28] = "Alcohol Usage"
nhanes_df_final$`Health Insurance Coverage`<- factor(nhanes_df_final$`Health Insurance Coverage`)

head(nhanes_df_final)
# make Blood Lead Level quartiles


set.seed(123) #to ensure same result throughout replication
nhanes_quantcut<- quantcut(nhanes_df_final$`Blood Lead Level`,q = 4)
table(nhanes_quantcut)
nhanes_quantcut_rank <-ifelse(nhanes_quantcut =="[0.05,0.88]", "Q1",
                               ifelse(nhanes_quantcut == "(0.88,1.4]", "Q2",
                                      ifelse(nhanes_quantcut =="(1.4,2.23]","Q3", "Q4")))
table(nhanes_quantcut_rank)
nhanes_df_final$BloodLeadQuartile<- nhanes_quantcut_rank

# # ok super happy fun analysis time!!

```
```{r}

og_table3 <- CreateTableOne(data = nhanes_df_final,
vars = c("Blood Lead Level", "Average Systolic Blood Pressure",
"Average Diastolic Blood Pressure","Education level", "Age", "Race/Ethnicity",
"Ratio of Family Income to Poverty Level","BMI", "Alcohol Usage", "Smoking Status",
"Hypertension Status", "Sex", "Health Insurance Coverage"), strata = "Race/Ethnicity")
kableone(og_table3)

```

```{r}

# Make demo tables and add BLL quartile
# stratification
og_table1 <- CreateTableOne(data = nhanes_df_final,
vars = c("Blood Lead Level", "Average Systolic Blood Pressure",
"Average Diastolic Blood Pressure",
"Education level", "Age", "Race/Ethnicity","Ratio of Family Income to Poverty Level",
"BMI", "Alcohol Usage", "Smoking Status",
"Hypertension Status", "Sex", "Health Insurance Coverage"))


kableone(og_table1)





og_table2 <- CreateTableOne(data = nhanes_df_final,
vars = c("Blood Lead Level", "Average Systolic Blood Pressure",
"Average Diastolic Blood Pressure","Education level", "Age", "Race/Ethnicity",
"Ratio of Family Income to Poverty Level","BMI", "Alcohol Usage", "Smoking Status",
"Hypertension Status", "Sex", "Health Insurance Coverage"), strata = "BloodLeadQuartile")
kableone(og_table2)



#save data
#save(nhanes_df_final, file = "NHANES_clean.RData")
# 
# #Load 
#load("NHANES_clean.RData")
```

The table shows various parameters measured in 4 quantiles (Q1, Q2, Q3, Q4) and their corresponding statistics (mean and standard deviation (SD)) for different groups or categories, as well as p-values.
In summary, these results indicate strong associations between the listed factors (Blood Lead Level, Blood Pressure, Education, Age, Race/Ethnicity, Income, BMI, Alcohol Usage, Smoking Status, Hypertension Status, Sex, and Health Insurance Coverage) and the outcome being studied. The p-values less than 0.001 for all these factors suggest that they are statistically significant in their influence on the outcome.






