---
title: 'Code for Thesis: Overall'
output:
  pdfdocut: default
  html_docut: default
date: "2023-11-02"
---
Load libraries in which we will use in this analysis 
```{r}
library(dplyr)
library(gtsummary)
library(tidyverse)
library(tidyselect)
library(kableExtra)
library(naniar)
library(ggplot2)
library(tidyr)
library(broom)
library(tableone)
library(stargazer)
library(magrittr)
library(GGally)
library(broom)
library(corrplot)

knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
#Load the pre-processed data 
load("C:/Users/CAU Student/Documents/VG Thesis/NHANES_clean.RData")

new_df<-nhanes_df_final%>%
  dplyr::select(`Blood Lead Level`, `Hypertension Diagnosis`, `Hypertension Status`,BPQ040A,BPXSY1,BPXSY2,BPXSY3,BPXSY4,BPXDI1,BPXDI2,BPXDI3,BPXDI4,Age,`Race/Ethnicity`, `Ratio of Family Income to Poverty Level`,`Education level`, Sex,BMI, `Number of Drinks/Year`,SMQ040,SMQ020, `Health Insurance Coverage`,`Average Systolic Blood Pressure`,`Average Diastolic Blood Pressure`, `Smoking Status`, `Alcohol Usage`,BloodLeadQuartile)%>%
  dplyr::rename(BLL=`Blood Lead Level`, hypertension_diagnosis=`Hypertension Diagnosis`, hypertension_status=`Hypertension Status`,age=Age,race=`Race/Ethnicity`, income=`Ratio of Family Income to Poverty Level`,education=`Education level`, sex=Sex, num_drinks=`Number of Drinks/Year`, HIC=`Health Insurance Coverage`,systolic_bp=`Average Systolic Blood Pressure`, diastolic_bp= `Average Diastolic Blood Pressure`, smoking=`Smoking Status`, alcohol=`Alcohol Usage`)


# #average_bll_by_race <- new_df %>%
#   select(race,BLL)%>%
#   group_by(race)%>%
#   summarise(average_bll = mean(BLL, na.rm = TRUE))
# 
# # Print the result
# #print(average_bll_by_race)
# 
# # Calculate average BLL by race using aggregate
# average_bll_by_race <- aggregate(BLL ~ race, data = new_df, FUN = mean, na.rm = TRUE)

# Print the result
#print(average_bll_by_race)


```
### Plots for 
```{r}
# Create the plots

new_df$educ_recode <- ifelse(new_df$education == 1, "> 9th grade", 
                              ifelse(new_df$education == 2, "HS & no diploma", 
                              ifelse(new_df$education == 3, "HS graduate/GED", 
                              ifelse(new_df$education == 4, "Some college or AA degree", 
                              ifelse(new_df$education == 5, "college graduate or higher",
                              ifelse(new_df$education == 7, "Don't know",
                              ifelse(new_df$education == 9, "Refused",NA)))))))

new_df$educ_recode <- factor(new_df$educ_recode, 
                              levels = c("> 9th grade", "HS & no diploma", 
                                         "HS graduate/GED", "Some college or AA degree", 
                                         "college graduate or higher", "Don't know", 
                                         "Refused"))


# sys_educ<-ggplot(new_df, aes(x = educ_recode, y = systolic_bp, fill = educ_recode)) +
#   geom_boxplot() +
#   labs(title = "Systolic Blood Pressure by Education Category",
#        x = "Education",
#        y = "Systolic Blood Pressure (mmHg)",
#        fill = "Education") +
#   theme_minimal()+
#   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
# sys_educ



sys_educ <- ggplot(new_df, aes(x = educ_recode, y = systolic_bp, fill = educ_recode)) +
  geom_boxplot() +
  labs(title = "Systolic Blood Pressure by Education Category",
       x = NULL,  # Removing x-axis label
       y = "Systolic Blood Pressure (mmHg)",
       fill = "Education") +
  theme_minimal() +
  theme(axis.text.x = element_blank())  # Removing x-axis labels

sys_educ

new_df$race_recode<- ifelse(new_df$race == 1, "Mexican American", 
                              ifelse(new_df$race == 2, "Other Hispanic", 
                              ifelse(new_df$race == 3, "NH White", 
                              ifelse(new_df$race == 4, "NH Black", 
                              ifelse(new_df$race == 5, "NH Multiracial", NA)))))

# Create boxplot with recoded race variable
sys_race <- ggplot(new_df, aes(x = race_recode, y = systolic_bp, fill = race_recode)) +
  geom_boxplot() +
  labs(title = "Systolic Blood Pressure by Race/Ethnicity",
       x = "Race/Ethnicity",
       y = "Systolic Blood Pressure (mmHg)",
       fill = "Race/Ethnicity") +
  theme_minimal()

sys_race

# Calculate standard errors
se_bll<- aggregate(BLL ~ race_recode, data = new_df, FUN = function(x) sd(x)/sqrt(length(x)))

bll_race <- ggplot(new_df, aes(x = race_recode, y = BLL, fill = race_recode)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_errorbar(data = se_bll, aes(ymin = BLL - BLL, ymax = BLL + BLL, group = race_recode),
                position = position_dodge(width = 0.9), width = 0.25) +
  labs(title = "Average Blood Lead Level by Race Category",
       x = "Race/Ethnicity",
       y = "Average BLL",
       fill = "Race") +
  theme_minimal()

bll_race


# Create scatter plot
scatter_plot<- ggplot(new_df, aes(x = BLL, y = systolic_bp, color = race_recode)) +
  geom_point() +
  labs(title = "Scatter Plot of Blood Lead Level and Systolic Blood Pressure by Race",
       x = "Blood Lead Level (BLL)",
       y = "Systolic Blood Pressure (mmHg)",
       color = "Race") +
  theme_minimal()

scatter_plot

# Calculate average BLL and systolic blood pressure for each race
avg_data<- aggregate(systolic_bp ~ race+BLL, data = new_df, FUN = mean)

# Create line plot
 line_plot<- ggplot(avg_data, aes(x = BLL, y = systolic_bp, color = race)) +
  geom_line(linewidth = 1.5) +  # Add lines
  labs(title = "Average Systolic Blood Pressure by Blood Lead Level and Race",
       x = "Blood Lead Level (BLL)",
       y = "Systolic Blood Pressure (mmHg)",
      color = "Race") +
  theme_minimal()
#
# line_plot


 plt1<-ggpairs(new_df%>% select(BLL,systolic_bp,income,BMI)%>%na.omit(),
lower = list(continuous = wrap("smooth", alpha = 0.7, size=0.5)))+
 theme_bw()
 plt1

```

```{r}
#model 1 from paper definition
model1_systolic_paper<- lm(systolic_bp~BLL+race+sex, data = new_df)

#model2 from paper definition
model2_systolic_paper<- lm(systolic_bp~BLL+race+age+sex, data = new_df)

#model3 from paper definition
model3_systolic_paper<- lm(systolic_bp~BLL+race+age+education+income+BMI+alcohol+smoking+sex, data = new_df)

```

# Define a list of models

```{r}
model1_systolic<-lm(systolic_bp ~ BLL+race+sex,  data = new_df)
model2_systolic<-lm(systolic_bp~ BLL*race+age+sex,  data = new_df)
model3_systolic<-lm(systolic_bp ~ BLL+race+education+age+sex,  data = new_df)
```
# Perform Model Selection

```{r}
#Model selection tests
anova(model1_systolic)
AIC(model1_systolic)
BIC(model1_systolic)


anova(model2_systolic)
AIC(model2_systolic)
BIC(model2_systolic)


anova(model3_systolic)
AIC(model3_systolic)
BIC(model3_systolic)


#Create a contingency table
contingency_table<- table(new_df$race, new_df$education)

# Perform chi-square test
chi_square_test<- chisq.test(contingency_table)

# Print the results
print(chi_square_test)


# Perform t-test
t_test_result_educ<- t.test(systolic_bp ~ hypertension_status, data = new_df)

# Print the results
print(t_test_result_educ)


# Perform Pearson correlation coefficient test
cor_test_result<- cor.test(new_df$systolic_bp, new_df$BLL, method = "pearson")

# Print the results
print(cor_test_result)
```
# Model 1
#PAPER RESULTS
## Dummy Coding (coefficients to replicate paper)

```{r}
# Create an empty list to store models
models <- list()

# Fit and store each model in the list
models$model1_systolic_pap_dummy<- lm(systolic_bp ~ BLL, data = new_df)

# Create an empty list to store model summaries
model_summaries <- list()

# Loop over sex and race levels
for (x in unique(new_df$sex)){
  for (z in unique(new_df$race)) {
    # Subset the data for the current race and sex category
    subset_data <- subset(new_df, race == z & sex==x)
    
    # Fit Model 1 on the subset data
    model_result <- tidy(lm(systolic_bp ~ BLL, data = subset_data))
    
    # Extract coefficients
    coef <- model_result$estimate[model_result$term == "BLL"]
    
    # Extract confidence intervals
    conf_int <- confint(lm(systolic_bp ~ BLL, data = subset_data))["BLL", ]
    
    #Extract p-values
    p_value <- coef(summary(lm(systolic_bp ~ BLL, data = subset_data)))["BLL", "Pr(>|t|)"]

    
    # Create a data frame with the results
    summary <- data.frame(
      model = "model1_systolic",
      race = z,
      sex=x,
      coef = coef,
      conf_int = conf_int,
      p_value=p_value
    )
    model_summaries <- c(model_summaries, list(summary))
  }
}

# Print the resulting data frame
model_summaries_df <- do.call(rbind, model_summaries)
# Now you have a data frame model_summaries_df with summaries stratified by sex and race for Model 1
print(model_summaries_df)

```

```{r}

# Create an empty list to store models
models2<- list()

# Fit and store each model in the list
models2$model2_systolic_pap_dummy<- lm(systolic_bp ~ BLL+age, data = new_df)

# Create an empty list to store model summaries
model_summaries2<- list()

# Loop over sex and race levels
for (x in unique(new_df$sex)){
  for (z in unique(new_df$race)) {
    # Subset the data for the current race and sex category
    subset_data2<- subset(new_df, race == z & sex==x)
    
    # Fit Model 1 on the subset data
    model_result2<- tidy(lm(systolic_bp ~ BLL+age, data = subset_data2))
    
    # Extract coefficients
    coef <- model_result2$estimate[model_result2$term == "BLL"]
    
    # Extract confidence intervals
    conf_int <- confint(lm(systolic_bp ~ BLL+age, data = subset_data2))["BLL", ]
    
    # Create a data frame with the results
    summary2<- data.frame(
      model = "model2_systolic",
      race = z,
      sex=x,
      coef = coef,
      conf_int = conf_int
    )
    model_summaries2<- c(model_summaries2, list(summary2))
  }
}

# Print the resulting data frame
model_summaries_df2<- do.call(rbind, model_summaries2)
# Now you have a data frame model_summaries_df with summaries stratified by sex and race for Model 1
print(model_summaries_df2)

```

```{r}

# Create an empty list to store models
models3<- list()

# Fit and store each model in the list
models3$model3_systolic_pap_dummy<- lm(systolic_bp ~ BLL+age+education +income +BMI +alcohol +smoking, data= new_df)

# Create an empty list to store model summaries
model_summaries3<-list()

# Loop over sex and race levels
for (x in unique(new_df$sex)){
  for (z in unique(new_df$race)) {
    # Subset the data for the current race and sex category
    subset_data3<- subset(new_df, race == z & sex==x)
    
    # Fit Model 1 on the subset data
    model_result3<- tidy(lm(systolic_bp ~ BLL+age+education+income +BMI +alcohol +smoking, data = subset_data3))
    
    # Extract coefficients
    coef<- model_result3$estimate[model_result3$term == "BLL"]
    
    # Extract confidence intervals
    conf_int<- confint(lm(systolic_bp ~ BLL+age+education+income +BMI +alcohol+smoking, data = subset_data3))["BLL", ]
    
    # Create a data frame with the results
    summary3<- data.frame(
      model = "model3_systolic",
      sex=x,
      race = z,
      coef = coef,
      conf_int = conf_int
    )
    model_summaries3<- c(model_summaries3, list(summary3))
  }
}

# Print the resulting data frame
model_summaries_df3<- do.call(rbind, model_summaries3)
# Now you have a data frame model_summaries_df with summaries stratified by sex and race for Model 1
print(model_summaries_df3)

```
# Model 1
```{r}
model1_systolic<-lm(systolic_bp ~ BLL+race+sex,  data = new_df)
summary(model1_systolic)
#Calculate confidence intervals for coefficients
conf_intervals_model1<- confint(model1_systolic)
# Create a summary dataframe with coefficients and confidence intervals
summary_dfmodel1<- cbind(coefficients = coef(model1_systolic), conf_intervals_model1)
# Print the summary dataframe
print(summary_dfmodel1)
```

## Simple Coding
```{r}
# Function to create simple regression coding matrix
create_regression_coding <- function(k) {
  # Initialize coding matrix
  coding_matrix <- matrix(0, nrow = k, ncol = k - 1)
  
  # Fill coding matrix with appropriate values
  for (i in 1:k) {
    if (i == 1) {
      coding_matrix[i, ] <- -1 / k
    } else {
      coding_matrix[i, i - 1] <- (k - 1) / k
    }
  }
  
  # Replace 0s with -1/k
  coding_matrix[coding_matrix == 0] <- -1 / k
  
  return(coding_matrix)
}

# Define the number of levels (k)
k <- 5

# Create regression coding matrix
regression_coding<-create_regression_coding(k)


# Print the regression coding matrix
print(regression_coding)


# Apply regression coding to the race variable in your dataframe
contrasts(new_df$race)<-regression_coding

# Print the summary of the model
summary(model1_systolic)
conf_intervals_dummy<- confint(model1_systolic)
# Create a summary dataframe with coefficients and confidence intervals
summary_dummy_model1<- cbind(coefficients = coef(model1_systolic), conf_intervals_dummy)
# Print the summary dataframe
print(summary_dummy_model1)


```
#Forward coding

```{r}
# Given matrix
forward_matrix<- matrix(c(0.8, 0.6, 0.4, 0.2,
                            -0.2, 0.6, 0.4, 0.2,
                            -0.2, -0.4, 0.4, 0.2,
                            -0.2, -0.4, -0.6, 0.2,
                            -0.2, -0.4, -0.6, -0.8), nrow = 5, ncol = 4, byrow = TRUE)


contrasts(new_df$race)<-forward_matrix
# Fit the model with the specified contrast coding
model1_systolic<-lm(systolic_bp ~ BLL+race+sex,  data = new_df)

# Print the model summary
summary_fwd_model1<-summary(model1_systolic)
# Adding 95% Confidence Intervals
conf_int_fwd_model1<- confint(model1_systolic)

# Combine the summary statistics and confidence intervals
summary_int_fwd_model1<- cbind(summary_fwd_model1$coefficients, conf_int_fwd_model1)

# Print the combined summary with confidence intervals
print(summary_int_fwd_model1)
```
#Backward Coding
```{r}
# Given matrix
backward_matrix <- matrix(c(-0.8, -0.6, -0.4, -0.2,
                            0.2, -0.6, -0.4, -0.2,
                            0.2, 0.4, -0.4, -0.2,
                            0.2, 0.4, 0.6, -0.2,
                            0.2, 0.4, 0.6, 0.8), nrow = 5, ncol = 4, byrow = TRUE)


contrasts(new_df$race)<-backward_matrix
# Fit the model with the specified contrast coding
model1_systolic<-lm(systolic_bp ~ BLL+race+sex,  data = new_df)

# Print the model summary
summary_bckwd_model1<-summary(model1_systolic)
# Adding 95% Confidence Intervals
conf_int_bckwd_model1<- confint(model1_systolic)

# Combine the summary statistics and confidence intervals
summary_int_bckwd_model1<- cbind(summary_bckwd_model1$coefficients, conf_int_bckwd_model1)

# Print the combined summary with confidence intervals
print(summary_int_bckwd_model1)
```
#Deviation Coding
```{r}
# Function to create a deviation coding matrix
create_deviation_coding <- function(k) {
  # Define the deviation coding matrix
  deviation_matrix <- matrix(0, nrow = k, ncol = k - 1)
  
  # Fill the matrix with appropriate values
  for (i in 1:k) {
    if (i == 1) {
      deviation_matrix[i, ] <- rep(-1, k - 1)
    } else {
      deviation_matrix[i, i - 1] <- 1
    }
  }
  
  return(deviation_matrix)
}

# Define the number of categories (k)
k <- 5  # Change this value as needed

# Create the deviation coding matrix
deviation_matrix <- create_deviation_coding(k)

# Assign row names
row.names(deviation_matrix) <- c("1", "2", "3", "4", "5")

# Add zeros for the additional column(s) in the matrix
deviation_matrix <- cbind(deviation_matrix, matrix(0, nrow = k, ncol = k - nrow(deviation_matrix)))

# Apply deviation coding to the race variable
contrasts(new_df$race)<- deviation_matrix

model1_systolic<-lm(systolic_bp ~ BLL+race+sex,  data = new_df)

# Print the model summary
summary_dev_model1<-summary(model1_systolic)
# Adding 95% Confidence Intervals
conf_int_dev_model1<- confint(model1_systolic)

# Combine the summary statistics and confidence intervals
summary_int_dev_model1<- cbind(summary_dev_model1$coefficients, conf_int_dev_model1)

# Print the combined summary with confidence intervals
print(summary_int_dev_model1)
```

# Model 2

```{r}
model2_systolic<-lm(systolic_bp~ BLL*race+age+sex,  data = new_df)
summary(model2_systolic)
conf_intervals_model2<- confint(model2_systolic)
# Create a summary dataframe with coefficients and confidence intervals
summary_df_model2<- cbind(coefficients = coef(model2_systolic), conf_intervals_model2)
# Print the summary dataframe
print(summary_df_model2)
```
## Simple Coding
```{r}
# Function to create simple regression coding matrix
create_regression_coding <- function(k) {
  # Initialize coding matrix
  coding_matrix <- matrix(0, nrow = k, ncol = k - 1)
  
  # Fill coding matrix with appropriate values
  for (i in 1:k) {
    if (i == 1) {
      coding_matrix[i, ] <- -1 / k
    } else {
      coding_matrix[i, i - 1] <- (k - 1) / k
    }
  }
  
  # Replace 0s with -1/k
  coding_matrix[coding_matrix == 0] <- -1 / k
  
  return(coding_matrix)
}

# Define the number of levels (k)
k <- 5

# Create regression coding matrix
regression_coding <- create_regression_coding(k)


# Print the regression coding matrix
print(regression_coding)

# Assuming you already have the regression coding matrix stored in 'regression_coding'

# Apply regression coding to the race variable in your dataframe
contrasts(new_df$race)<- regression_coding

# Print the summary of the model
summary(model2_systolic)
conf_intervals_dummy_model2<- confint(model2_systolic)
# Create a summary dataframe with coefficients and confidence intervals
summary_dummy_model2<- cbind(coefficients = coef(model2_systolic), conf_intervals_dummy_model2)
# Print the summary dataframe
print(summary_dummy_model2)


```
#Forward coding

```{r}
# Given matrix
forward_matrix<- matrix(c(0.8, 0.6, 0.4, 0.2,
                            -0.2, 0.6, 0.4, 0.2,
                            -0.2, -0.4, 0.4, 0.2,
                            -0.2, -0.4, -0.6, 0.2,
                            -0.2, -0.4, -0.6, -0.8), nrow = 5, ncol = 4, byrow = TRUE)


contrasts(new_df$race)<-forward_matrix
# Fit the model with the specified contrast coding
model2_systolic<-lm(systolic_bp~ BLL*race+age+sex,  data = new_df)

# Print the model summary
summary_fwd_model2<-summary(model2_systolic)
# Adding 95% Confidence Intervals
conf_int_fwd_model2<- confint(model2_systolic)

# Combine the summary statistics and confidence intervals
summary_int_fwd_model2<- cbind(summary_fwd_model2$coefficients, conf_int_fwd_model2)

# Print the combined summary with confidence intervals
print(summary_int_fwd_model2)
```
#Backward Coding
```{r}
# Given matrix
backward_matrix <- matrix(c(-0.8, -0.6, -0.4, -0.2,
                            0.2, -0.6, -0.4, -0.2,
                            0.2, 0.4, -0.4, -0.2,
                            0.2, 0.4, 0.6, -0.2,
                            0.2, 0.4, 0.6, 0.8), nrow = 5, ncol = 4, byrow = TRUE)


contrasts(new_df$race)<-backward_matrix
# Fit the model with the specified contrast coding
model2_systolic<-lm(systolic_bp~ BLL*race+age+sex,  data = new_df)

# Print the model summary
summary_bckwd_model2<-summary(model2_systolic)
# Adding 95% Confidence Intervals
conf_int_bckwd_model2<- confint(model2_systolic)

# Combine the summary statistics and confidence intervals
summary_int_bckwd_model2<- cbind(summary_bckwd_model2$coefficients, conf_int_bckwd_model2)

# Print the combined summary with confidence intervals
print(summary_int_bckwd_model2)
```
#Deviation Coding
```{r}
# Function to create a deviation coding matrix
create_deviation_coding <- function(k) {
  # Define the deviation coding matrix
  deviation_matrix <- matrix(0, nrow = k, ncol = k - 1)
  
  # Fill the matrix with appropriate values
  for (i in 1:k) {
    if (i == 1) {
      deviation_matrix[i, ] <- rep(-1, k - 1)
    } else {
      deviation_matrix[i, i - 1] <- 1
    }
  }
  
  return(deviation_matrix)
}

# Define the number of categories (k)
k <- 5  # Change this value as needed

# Create the deviation coding matrix
deviation_matrix <- create_deviation_coding(k)

# Assign row names
row.names(deviation_matrix) <- c("1", "2", "3", "4", "5")

# Add zeros for the additional column(s) in the matrix
deviation_matrix <- cbind(deviation_matrix, matrix(0, nrow = k, ncol = k - nrow(deviation_matrix)))

# Apply deviation coding to the race variable
contrasts(new_df$race)<- deviation_matrix

model2_systolic<-lm(systolic_bp~ BLL*race+age+sex,  data = new_df)

# Print the model summary
summary_dev_model2<-summary(model2_systolic)
# Adding 95% Confidence Intervals
conf_int_dev_model2<- confint(model2_systolic)

# Combine the summary statistics and confidence intervals
summary_int_dev_model2<- cbind(summary_dev_model2$coefficients, conf_int_dev_model2)

# Print the combined summary with confidence intervals
print(summary_int_dev_model2)
```

# Model 3
```{r}
model3_systolic<-lm(systolic_bp ~ BLL+race+education+age+sex,  data = new_df)
summary(lm(systolic_bp~ BLL+race+education+age+sex, data=new_df))
conf_intervals_model3<- confint(model3_systolic)
# Create a summary dataframe with coefficients and confidence intervals
summary_dfmodel3<- cbind(coefficients = coef(model3_systolic), conf_intervals_model3)
# Print the summary dataframe
print(summary_dfmodel3)
```

## Simple Coding
```{r}
# Function to create simple regression coding matrix
create_regression_coding <- function(k) {
  # Initialize coding matrix
  coding_matrix <- matrix(0, nrow = k, ncol = k - 1)
  
  # Fill coding matrix with appropriate values
  for (i in 1:k) {
    if (i == 1) {
      coding_matrix[i, ] <- -1 / k
    } else {
      coding_matrix[i, i - 1] <- (k - 1) / k
    }
  }
  
  # Replace 0s with -1/k
  coding_matrix[coding_matrix == 0] <- -1 / k
  
  return(coding_matrix)
}

# Define the number of levels (k)
k <- 5

# Create regression coding matrix
regression_coding <- create_regression_coding(k)


# Print the regression coding matrix
print(regression_coding)

# Assuming you already have the regression coding matrix stored in 'regression_coding'

# Apply regression coding to the race variable in your dataframe
contrasts(new_df$race)<- regression_coding

# Print the summary of the model
summary(model3_systolic)
conf_intervals_dummy_model3<- confint(model3_systolic)
# Create a summary dataframe with coefficients and confidence intervals
summary_dummy_model3<- cbind(coefficients = coef(model3_systolic), conf_intervals_dummy_model3)
# Print the summary dataframe
print(summary_dummy_model3)


```
#Forward coding

```{r}
# Given matrix
forward_matrix<- matrix(c(0.8, 0.6, 0.4, 0.2,
                            -0.2, 0.6, 0.4, 0.2,
                            -0.2, -0.4, 0.4, 0.2,
                            -0.2, -0.4, -0.6, 0.2,
                            -0.2, -0.4, -0.6, -0.8), nrow = 5, ncol = 4, byrow = TRUE)


contrasts(new_df$race)<-forward_matrix
# Fit the model with the specified contrast coding
model3_systolic<-lm(systolic_bp ~ BLL+race+education+age+sex,  data = new_df)

# Print the model summary
summary_fwd_model3<-summary(model3_systolic)
# Adding 95% Confidence Intervals
conf_int_fwd_model3<- confint(model3_systolic)

# Combine the summary statistics and confidence intervals
summary_int_fwd_model3<- cbind(summary_fwd_model3$coefficients, conf_int_fwd_model3)

# Print the combined summary with confidence intervals
print(summary_int_fwd_model3)
```
#Backward Coding
```{r}
# Given matrix
backward_matrix <- matrix(c(-0.8, -0.6, -0.4, -0.2,
                            0.2, -0.6, -0.4, -0.2,
                            0.2, 0.4, -0.4, -0.2,
                            0.2, 0.4, 0.6, -0.2,
                            0.2, 0.4, 0.6, 0.8), nrow = 5, ncol = 4, byrow = TRUE)


contrasts(new_df$race)<-backward_matrix
# Fit the model with the specified contrast coding
model3_systolic<-lm(systolic_bp ~ BLL+race+education+age+sex,  data = new_df)

# Print the model summary
summary_bckwd_model3_<-summary(model3_systolic)
# Adding 95% Confidence Intervals
conf_int_bckwd_model3_ <- confint(model3_systolic)

# Combine the summary statistics and confidence intervals
summary_int_bckwd_model3_<- cbind(summary_bckwd_model3_$coefficients, conf_int_bckwd_model3_)

# Print the combined summary with confidence intervals
print(summary_int_bckwd_model3_)
```
#Deviation Coding
```{r}
# Function to create a deviation coding matrix
create_deviation_coding <- function(k) {
  # Define the deviation coding matrix
  deviation_matrix <- matrix(0, nrow = k, ncol = k - 1)
  
  # Fill the matrix with appropriate values
  for (i in 1:k) {
    if (i == 1) {
      deviation_matrix[i, ] <- rep(-1, k - 1)
    } else {
      deviation_matrix[i, i - 1] <- 1
    }
  }
  
  return(deviation_matrix)
}

# Define the number of categories (k)
k <- 5  # Change this value as needed

# Create the deviation coding matrix
deviation_matrix <- create_deviation_coding(k)

# Assign row names
row.names(deviation_matrix) <- c("1", "2", "3", "4", "5")

# Add zeros for the additional column(s) in the matrix
deviation_matrix <- cbind(deviation_matrix, matrix(0, nrow = k, ncol = k - nrow(deviation_matrix)))

# Apply deviation coding to the race variable
contrasts(new_df$race)<- deviation_matrix

model3_systolic<-lm(systolic_bp ~ BLL+race+education+age+sex,  data = new_df)

# Print the model summary
summary_dev_model3<-summary(model3_systolic)
# Adding 95% Confidence Intervals
conf_int_dev_model3<- confint(model3_systolic)

# Combine the summary statistics and confidence intervals
summary_int_dev_model3<- cbind(summary_dev_model3$coefficients, conf_int_dev_model3)

# Print the combined summary with confidence intervals
print(summary_int_dev_model3)
```
